Robust Control of a Fluxonium Qubit

### 2020-11-30 paper/4longitude.tex
updated figures, lots of proofreading

\begin{figure*}[ht]

  \begin{subfigure}{.315\textwidth}
    \includegraphics[width=\linewidth]{assets/f1a.png}
    \caption{\label{fig:longitudea}}
  \end{subfigure}\hfill
  \begin{subfigure}{.23\textwidth}
    \includegraphics[width=\linewidth]{assets/f1b.png}
    \caption{\label{fig:longitudeb}}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f1c.png}
    \caption{\label{fig:longitudec}}
  \end{subfigure}
  \caption{
    (a) Numerically optimized gates (red) and analytically optimized gates (blue).
    (b) $T_{1}$ interpolation function used in optimization. Markers
    denote the time-averaged, absolute amplitude of each gate.
    (c) Lindblad master equation simulation with $T_{1}$ dissipation
    for successive gate applications. The cumulative
    gate error is computed after each gate application.
    The numerically optimized $Z/2$ and $Y/2$ gate errors are indistinguishable.
  }
  \label{fig:longitude}
\end{figure*}

\section{Longitudinal Relaxation Awareness \label{sec:longitude}}
The strength of longitudinal relaxation type decoherence varies
with the control parameters in a range of superconducting circuit platforms,
including the fluxonium qubit. Because dissipation to the thermal bath
via longitudinal relaxation is an irreversible process that results in information loss,
it is advantageous to tune the controls to extend the longitudinal
relaxation time $T_{1}$--which represents the $1/e$ relaxation time
of the quantum state.
Computing the gate error due to longitudinal relaxation
requires propagating density matrices of size $n \times n$ under master equation
dynamics, rather than state vectors of size $n$ under the TDSE dynamics \eqref{eq:tdse}.
We avoid this increase in computational complexity by
penalizing the probability (integrated rate) of longitudinal relaxation.
Using this probability as proxy for the gate error incurred
is reasonable because losses due to longitudinal relaxation increase monotonically
in time.
This technique can be extended to
error channels which share the monotonically increasing property.
Additionally, for a constant $T_{1}$ time, a shorter gate duration
would favor a lower longitudinal relaxation probability. We allow
the optimizer to tune the gate duration in order to minimize the
longitudinal relaxation probability. Our scheme for time-optimal
control is applicable to any time-optimal problem, not only
the one we study here.

The longitudinal relaxation probability is given by,
\begin{equation}
  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
\end{equation}
This value is appended to the augmented state vector \eqref{eq:astatecontrols}
and penalized according to \eqref{eq:costfun}.
$T_{1}(a_{k})$ is obtained at each knot point by evaluating
a spline fit to experimental data of the form $\{(a, T_{1})\}$,
see Figure \ref{fig:longitudeb}.
It is also possible to fit a spline to theoretically obtained data.
However, $T_{1}$ values are known to fluctuate greatly
with laboratory temperatures \cite{klimov2018fluctuations}.
Interpolating $T_{1}$ from experimental data
increases the fridge truth of the simulation.

We allow the optimizer to tune the gate duration by
making the time step between each knot point $\Delta t_{k}$
a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
than the number of knot points $N$, preserves the
Markovian decision structure of the trajectory
optimization problem. The square root of the time step $\sqrt{\Delta t_{k}}$
is appended to the augmented control vector \eqref{eq:astatecontrols}
and the squared root
of the time step $\lvert \Delta t_{k} \rvert$ is used
for integration in the discrete dynamics function \eqref{eq:dyn_con}.
To ensure numerical
integration accuracy is maintained we add a bound
constraint on the time step at each knot point.

%% F2
\begin{figure*}[ht]
  \begin{subfigure}{.315\textwidth}
    \includegraphics[width=\linewidth]{assets/f2a.png}
    \caption{\label{fig:statica}}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f2c.png}
    \caption{\label{fig:staticb}}
  \end{subfigure}\hfill
  \begin{subfigure}{.23\textwidth}
    \includegraphics[width=\linewidth]{assets/f2b.png}
    \caption{\label{fig:staticc}}
  \end{subfigure}
  \caption{
    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
    analytic, sampling, unscented sampling, and the 1\textsuperscript{st}-
    and 2\textsuperscript{nd}-order derivative methods. The gates
    shown for the numerical methods are the solutions at twice the analytic
    gate time.
    (b) Single gate error as
    a function of the gate duration at a one-percent
    detuning from the nominal qubit frequency for all methods. Missing
    data points represent solutions with a gate error above $5 \cdot 10^{-5}$.
    (c) Single gate error as a function of the detuning from the nominal
    qubit frequency. The solutions for the analytic and
    1\textsuperscript{st}-order derivative methods are shown at multiples
    of the analytic gate time. The performance of the two methods is
    indistinguishable at the analytic gate time $18$ns.
  }
  \label{fig:static}
\end{figure*}

We compare the numerical method we have developed to the analytic gates
on the task of achieving low gate errors in the presence of longitudinal relaxation
for the $Z/2$, $Y/2$, and $X/2$ gates.
The numerically optimized gates converge on similar solutions, a periodic
waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitudea}.
They extend their gate times
beyond their analytic counterparts, trading longer gate times for access
to higher amplitudes and therefore higher $T_{1}$ times. All numerical gates reduce
their single gate errors by a factor of $5$ over
their analytic counterparts which is commensurate to their
probability of longitudinal relaxation reductions, see Appendix \ref{appendix:longitude}.
The gate error reported in this text is the infidelity
of the evolved state and the target state averaged over 1000 pseudo-randomly
generated initial states. The numerical $Z/2$ and $Y/2$ gates perform simlarly in
the concatenated gate application comparision, suppressing cumulative gate errors to $8 \cdot 10^{-3}$
over $2000$ gate applications $\sim 40 \mu\textrm{s}$, see Figure \ref{fig:longitudec}.
The numerical $X/2$ gate achieves a cumulative gate error of
$1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
These low cumulative gate errors for high gate counts are critical for
noisy, intermediate-scale quantum (NISQ) applications.
Both the analytic and numerical gates attain single gate errors sufficient for
quantum error correction ($< 10^{-4}$), which are required for fault-tolerant quantum computing
\cite{aharonov2008fault, fowler2009high, gottesman1997stabilizer}.
These improvements are significant for the realistic constraints we have imposed
on the gates, and do not represent a fundamental limit to the optimization methods we have
employed.

>>>>

%% F1
\begin{figure*}[ht]

  \begin{subfigure}{.315\textwidth}
    \includegraphics[width=\linewidth]{assets/f1a.png}
    \caption{\label{fig:longitudea}}
  \end{subfigure}\hfill
  \begin{subfigure}{.23\textwidth}
    \includegraphics[width=\linewidth]{assets/f1b.png}
    \caption{\label{fig:longitudeb}}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f1c.png}
    \caption{\label{fig:longitudec}}
  \end{subfigure}
  \caption{
    \todo{Ask Dave to okay figure color scheme}
    (a) Numerically optimized gates (dark) and analytically optimized gates (light).
    (b) $T_{1}$ interpolation function used in optimization. Markers
    denote the time-averaged, absolute amplitude of each gate.
    (c) Lindblad master equation simulation with $T_{1}$ dissipation
    for successive gate applications. The cumulative
    gate error is computed after each gate application.
    The numerically optimized $Z/2$ and $Y/2$ gate errors are indistinguishable.
  }
  \label{fig:longitude}
\end{figure*}

\section{Longitudinal Relaxation Awareness \label{sec:longitude}}
The strength of longitudinal relaxation type decoherence varies
with the control parameters in a range of superconducting circuit platforms,
including the fluxonium qubit. Because dissipation to the thermal bath
via longitudinal relaxation is an irreversible process that results in information loss,
it is advantageous to tune the controls to extend the longitudinal
relaxation time $T_{1}$--which represents the $1/e$ relaxation time
of the quantum state.
Typical approaches completely model the gate error due to longitudinal
relaxation by performing optimal control on $n^{2}$
density matrices of size $n \times n$ that evolve under master
equation dynamics, rather than $n$ states vectors of size $n$
under the TDSE \eqref{eq:tdse}.
We avoid this increase in computational complexity by
penalizing the probability (integrated rate) of longitudinal relaxation.
Using this probability as a proxy for the gate error incurred
is reasonable because losses due to longitudinal relaxation increase monotonically
in time.
This technique can be extended to
error channels which share the monotonically increasing property.
Additionally, for a constant $T_{1}$ time, a shorter gate time
would favor a lower longitudinal relaxation probability. We allow
the optimizer to tune the gate time to minimize the
longitudinal relaxation probability. Our scheme for time-optimal
control is suitable for any time-optimal problem, not only
the one we study here.

\todo{Ask Brian to review this paragraph and next}
The longitudinal relaxation probability is given by,
\begin{equation}
  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
\end{equation}
This value is appended to the augmented state vector \eqref{eq:astatecontrols}
and penalized according to \eqref{eq:costfun}.
$T_{1}(a_{k})$ is obtained at each knot point by evaluating
a spline fit to experimental data of the form $\{(a, T_{1})\}$,
see Figure \ref{fig:longitudeb}.
It is also possible to fit a spline to theoretically obtained data.
However, $T_{1}$ values are known to fluctuate greatly
with laboratory temperatures \cite{klimov2018fluctuations}.
Interpolating $T_{1}$ from experimental data
increases the fridge truth of the simulation.

We allow the optimizer to tune the gate time by
making the time step between each knot point $\Delta t_{k}$
a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
than the number of knot points $N$, preserves the
Markovian decision structure of the trajectory
optimization problem. The square root of the time step $\sqrt{\Delta t_{k}}$
is appended to the augmented control vector \eqref{eq:astatecontrols}
and the squared root
of the time step $\lvert \Delta t_{k} \rvert$ is used
for integration in the discrete dynamics function \eqref{eq:dyn_con}.
To ensure numerical
integration accuracy is maintained we add a bound
constraint on the time step at each knot point.

%% F2
\begin{figure*}[ht]
  \begin{subfigure}{.315\textwidth}
    \includegraphics[width=\linewidth]{assets/f2a.png}
    \caption{\label{fig:statica}}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f2b.png}
    \caption{\label{fig:staticb}}
  \end{subfigure}\hfill
  \begin{subfigure}{.23\textwidth}
    \includegraphics[width=\linewidth]{assets/f2c.png}
    \caption{\label{fig:staticc}}
  \end{subfigure}
  \caption{
    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
    analytic, sampling, unscented sampling, and the 1\textsuperscript{st}-
    and 2\textsuperscript{nd}-order derivative methods. The gates
    shown for the numerical methods are the solutions at twice the analytic
    gate time.
    (b) Single gate error as
    a function of the gate time at a one-percent
    qubit frequency detuning for all methods. Missing
    data points represent solutions with a gate error above $5 \cdot 10^{-5}$.
    (c) Single gate error as a function of the qubit frequency detuning.
    The solutions for the analytic and
    1\textsuperscript{st}-order derivative methods are shown at multiples
    of the analytic gate time. The performance for the two methods is
    indistinguishable at the analytic gate time $18$ns.
  }
  \label{fig:static}
\end{figure*}

We compare the numerical method we have developed to the analytic gates
on the task of achieving low gate errors in the presence of longitudinal relaxation
for the $Z/2$, $Y/2$, and $X/2$ gates.
The numerically optimized gates converge on similar solutions, a periodic
waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitudea}.
They extend their gate times
beyond their analytic counterparts, trading longer gate times for access
to higher amplitudes and therefore higher $T_{1}$ times. All numerical gates reduce
their single gate errors by a factor of $5$ over
their analytic counterparts, commensurate to their
probability of longitudinal relaxation reductions, see Appendix \ref{appendix:longitude}.
The gate error reported in this text is the infidelity
of the evolved state and the final state averaged over 1000 pseudo-randomly
generated initial states. The numerical $Z/2$ and $Y/2$ gates perform similarly in
the concatenated gate application comparison, suppressing cumulative gate errors to $8 \cdot 10^{-3}$
over $2000$ gate applications $\sim 40 \mu\textrm{s}$, see Figure \ref{fig:longitudec}.
The numerical $X/2$ gate achieves a cumulative gate error of
$1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
These low cumulative gate errors for high gate counts are critical for
noisy, intermediate-scale quantum (NISQ) applications.
Both the analytic and numerical gates attain single gate errors sufficient for
quantum error correction ($< 10^{-4}$), which are required for fault-tolerant quantum computing
\cite{aharonov2008fault, fowler2009high, gottesman1997stabilizer}.
These improvements are significant for the realistic constraints we have imposed
on the gates, and do not represent a fundamental limit to the optimization methods we have
employed.

### 2020-11-30 paper/5static.tex
updated figures, lots of proofreading

\section{Robustness to Static Parameter Deviations \label{sec:static}}
\todo{standardize gate time vs gate duration} We have formulated the quantum optimal control
problem as an open loop optimization problem, i.e.
feedback from the experiment is not incorporated in optimization.
However, the device typically deviates from the Hamiltonian we use in optimization,
leading to poor experimental performance. We combat errors
of this form using robust control techniques,
making the state evolution insensitive
to Hamiltonian parameter deviations. As an example,
we mitigate errors arising from the drift and finite measurement
precision of the qubit frequency which modifies \eqref{eq:hamiltonian}
by $f_{q} \gets f_{q} + \delta f_{q}$.
We consider three robust control techniques:
the sampling method, the unscented sampling method,
and the derivative method.

The sampling method simultaneously optimizes over copies of a state,
each of which evolves with a distinct deviant parameter value.
A pair of two sample states $\psi_{\pm}$
is appended to the augmented state vector \eqref{eq:astatecontrols}
. The samples are initialized in the same state.
Four pairs
are appended in total corresponding to the initial states
$\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
(\ket{0} - \ket{1}) / \sqrt{2}\}$.
The initial states are chosen
so that their outer products span the operators on the
Hilbert space  \cite{chow2009randomized}.
The discrete dynamics \eqref{eq:dyn_con} is modified so that the
samples evolve under the Hamiltonian given in \eqref{eq:hamiltonian}
with $f_{q} \gets f_{q} \pm \sigma_{f_{q}}$ for a fixed $\sigma_{f_{q}}$.
The final state
for each sample is the image of its initial state
under the desired gate. Deviations from the final state are
penalized in the objective \eqref{eq:costfun}.

For the unscented sampling method, $2(2n + d)$ samples
are appended to the augmented state vector \eqref{eq:astatecontrols}
for each initial state from the operator basis. Here $2n = 4$ is twice the
dimension of the Hilbert space, resulting from the isomorphism \eqref{eq:isomorphism},
and $d = 1$ is the number of deviant parameters. The samples
are used to represent a Gaussian distribution over all $2n$
elements of the initial state, modeling
the uncertainty in the state as a result of the uncertainty in
the deviant parameter. Each sample evolves under the Hamiltonian
given in \eqref{eq:hamiltonian} with $f_{q} \gets f_{q} + \delta f_{q}$.
The qubit frequency detuning $\delta f_{q}$ is determined by the statistics
of the ensemble of samples at each knot point.
The discrete dynamics \eqref{eq:dyn_con} is modified to 
evolve the samples under the deviant Hamiltonian
and then apply the unscented transformation to the ensemble,
accurately preserving the first and second moments
of the distribution.
As in the sampling method,
deviations from the final state are penalized according to \eqref{eq:costfun}.
A detailed procedure for the unscented transformation is given
in Appendix \ref{appendix:unscented}.

The derivative method penalizes the sensitivty of the state
to the deviant parameter, which is encoded in the ($l$\textsuperscript{th}-order)
derivative $\partial_{f_{q}}^{l} \psi$. In the $m$\textsuperscript{th}-order
derivative method, all state derivatives of order $1, \dots, m$
are appended to the augmented state vector \eqref{eq:astatecontrols}
for each initial state in the operator basis.
The norms of the state derivatives are penalized
in \eqref{eq:costfun} by setting the corresponding final states to zero.
The state derivatives could be obtained in the discrete dynamics \eqref{eq:dyn_con}
with backwards mode differentiation.
Naive automatic differentiation
would differentiate all dynamics at knot points
$1, \dots, k - 1$ to obtain the state derivative at knot point $k$, requiring
$O(N^{2})$ matrix multiplications. Instead, we 
employ forwards mode differentiation on the TDSE to obtain coupled dynamics
which require $O(N)$ matrix multiplications to integrate.
For example, the dynamics for the $1$\textsuperscript{st}-order derivative method are,
\begin{align}
  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
  i \hbar \frac{d}{dt} \ket{\partial_{f_{q}}\psi} &=
  H \ket{\partial_{f_{q}} \psi} +
  (\partial_{f_{q}} H) \ket{\psi}
  \label{eq:d1dyn}
\end{align}
Exponential integrators that account for the non-linear
term in \eqref{eq:d1dyn} may be used to efficiently integrate the coupled dynamics,
see Appendix \ref{appendix:derivative}. For timing results
and an analysis of how the three robust methods scale to larger Hilbert spaces
and more deviant parameters,
consult Appendix \ref{appendix:time}.

We task these methods with achieving
a $Z/2$ gate subject to a static qubit frequency detuning.
We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
the sampling methods accordingly. We compute gate errors for each method
by evolving the optimized trajectory under the Hamiltonian in \eqref{eq:hamiltonain}
with $f_{q} \gets f_{q} + \delta f_{q}$ at the stated qubit frequency detuning.

We compare the numerical methods
to an analytically derived $Z/2$ gate, see Figure \ref{fig:statica}. 
The analytic gate corresponds to
idling at the flux frustration point $a = 0$. It
is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
is simple to derive. Its erroneous rotation angle $2 \pi t_{Z/2} \delta f_{q}$ is linearly sensitive to
the qubit frequency detuning, resulting in a gate error that is quadratically sensitive
to the qubit frequency detutning.
At a one-percent
qubit frequency detuning ($\delta f_{q} / f_{q} = 1\%$)
the analytic gate achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
which is sufficient for quantum error correction.
Although the analytic $Z/2$ gate performs well, it
only works at the gate time $t_{Z/2}$. The ability to perform $Z$ rotations in arbitrary times is critical
for multi-qubit experiments, where the qubits operate at different frequencies $f_{q, i} \neq f_{q, j}$.
Each numerical method is able to find solutions at
all gate times above $t_{Z/2}$, see Figure \ref{fig:staticb}.
These numerical methods offer an effective scheme for synchronizing
multi-qubit experiments.

The sampling and unscented sampling methods combine idling periods
with time-anti-symmetric ramps.
The gate error at $\delta f_{q} / f_{q} = 1\%$ for the sampling method
does not improve substantially over the
range of gate durations. The unscented sampling method
achieves linear decreases in its gate error at $\delta f_{q} / f_{q} = 1\%$ with longer gate durations
until half the larmor period $1 / 2 f_{q}$ after which it achieves a consistent
gate error $\sim 3.5 \cdot 10^{-5}$.

The derivative methods converge on qualitatively similar solutions that
use fast traingle pulses at the boundaries and balance time
on either side of the flux-frustration point symmetrically at low amplitudes.
Both methods achieve a super-linear scaling in their gate error as
a function of the gate duration. The gate error for the 1\textsuperscript{st}-order
derivative method approaches zero at the larmor period $1 / f_{q}$, see Figure \ref{fig:staticc}.
We believe the 1\textsuperscript{st}-order method outperforms the 2\textsuperscript{nd}-order
method due to the low contribution of second-order
terms to the gate error in this deviation regime, see Appendix \ref{appendix:derivative}.
The vanishing gate error at longer gate durations exhibited by the derivative methods
is reminiscent of the
ability of composite pulses to mitigate parameter deviation errors to arbitrary
order with sufficiently many pulses \cite{merrill2014progress}.
It is difficult to choose an appropriate composite pulse
for the problem studied here due to our Hamiltonian and experimental constraints.
We propose comparisons between composite pulses and numerical techniques
for future work.

%% F3
\begin{figure*}[ht]
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f3a.png}
    \caption{}
    \label{fig:stochastica}
  \end{subfigure}\hspace{0.025\textwidth}
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f3b.png}
        \caption{}
    \label{fig:stochasticb}
  \end{subfigure}
  \caption{
    \todo{fix the size of the images and where are x axis labels?}
    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
    sampling, unscented sampling, and the 1\textsuperscript{st}-
    and 2\textsuperscript{nd}-order derivative methods. The gates shown
    are the solutions at the analytic gate time.
    (b) Simulation of stochastic 1/$f$ flux noise for
    successive gate applications. The cumulative
    gate error is computed after each gate application.
  }
  \label{fig:stochastic}
\end{figure*}

>>>>

\section{Robustness to Static Parameter Deviations \label{sec:static}}
We have formulated the QOC
problem as an open-loop optimization problem, i.e.
we do not incorporate feedback from the experiment in optimization.
However, the device typically deviates from the Hamiltonian we use in optimization,
leading to poor experimental performance. We combat errors
of this form using robust control techniques,
making the state evolution insensitive
to Hamiltonian parameter deviations. As an example,
we mitigate errors arising from the drift and finite measurement
precision of the qubit frequency which modifies \eqref{eq:hamiltonian}
by $f_{q} \gets f_{q} + \delta f_{q}$.
We consider three robust control techniques:
the sampling method, the unscented sampling method,
and the derivative method.

\todo{Ask Brian to review this paragraph and next two}
The sampling method simultaneously optimizes over copies of a state,
each of which evolves with a distinct deviant parameter value.
We append four sets of two sample states $\psi_{\pm}$
to the augmented state vector \eqref{eq:astatecontrols}.
We initialize each pair of samples in the same state.
The four initial states are chosen so that their outerproducts
span the operators on the Hilbert space
$\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
The discrete dynamics \eqref{eq:dyn_con} is modified so that the
samples evolve under the Hamiltonian given in \eqref{eq:hamiltonian}
with $f_{q} \gets f_{q} \pm \sigma_{f_{q}}$ for a fixed $\sigma_{f_{q}}$.
The final state
for each sample is the image of its initial state
under the desired gate. We penalize deviations
from the final state in the objective \eqref{eq:costfun}.

For the unscented sampling method, we append $2(2n + d)$ samples
to the augmented state vector \eqref{eq:astatecontrols}
for each initial state in the operator basis. Here $2n = 4$ is twice the
dimension of the Hilbert space, resulting from the isomorphism \eqref{eq:isomorphism},
and $d = 1$ is the number of deviant parameters. The samples
represent a Gaussian distribution over the $2n$
elements of the initial state, modeling
the uncertainty in the state as a result of the uncertainty in
the deviant parameter. Each sample evolves under the Hamiltonian
given in \eqref{eq:hamiltonian} with $f_{q} \gets f_{q} + \delta f_{q}$.
We determine the qubit frequency detuning $\delta f_{q}$
at each knot point using the statistics of the samples.
We modify the discrete dynamics \eqref{eq:dyn_con} to
evolve the samples under the deviant Hamiltonian
and then apply the unscented transformation to the ensemble
of samples,
accurately preserving the first and second moments
of the distribution.
As in the sampling method,
we penalize deviations from the final state according to \eqref{eq:costfun}.
A detailed procedure for the unscented transformation is given
in Appendix \ref{appendix:unscented}.

The derivative method penalizes the sensitivty of the state
to the deviant parameter, which is encoded in the ($l$\textsuperscript{th}-order)
derivative $\partial_{f_{q}}^{l} \psi$. In the $m$\textsuperscript{th}-order
derivative method, we append all state derivatives of order $1, \dots, m$
to the augmented state vector \eqref{eq:astatecontrols}
for each initial state in the operator basis.
We penalize the norms of the state derivatives
in \eqref{eq:costfun} by setting the corresponding final states to zero.
We could obtain the state derivatives in the discrete dynamics \eqref{eq:dyn_con}
with backwards mode differentiation.
Naive automatic differentiation 
would differentiate all dynamics at knot points
$1, \dots, k - 1$ to obtain the state derivative at knot point $k$, requiring
$O(N^{2})$ matrix multiplications. Instead, we 
employ forwards mode differentiation on the TDSE to obtain coupled dynamics
which require $O(N)$ matrix multiplications to integrate.
For example, the dynamics for the $1$\textsuperscript{st}-order derivative method are:
\begin{align}
  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
  i \hbar \frac{d}{dt} \ket{\partial_{f_{q}}\psi} &=
  H \ket{\partial_{f_{q}} \psi} +
  (\partial_{f_{q}} H) \ket{\psi}
  \label{eq:d1dyn}
\end{align}
Exponential integrators that account for the non-linear
term in \eqref{eq:d1dyn} are used to efficiently integrate the coupled dynamics,
see Appendix \ref{appendix:derivative}. For timing
results and complexity analyses for the three methods,
consult Appendix \ref{appendix:time}.

We challenge these methods to achieve
a $Z/2$ gate subject to a static qubit frequency detuning.
We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
the sampling methods accordingly. We compute gate errors for each method
by evolving the system under the Hamiltonian in \eqref{eq:hamiltonian}
with the optimized flux amplitude and $f_{q} \gets f_{q} + \delta f_{q}$
at the stated qubit frequency detuning.

We compare the numerical methods
to an analytically derived $Z/2$ gate, see Figure \ref{fig:statica}. 
The analytic gate corresponds to
idling at the flux frustration point $a = 0$. It
is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
is simple to derive. Its erroneous rotation angle
$2 \pi t_{Z/2} \delta f_{q}$ is linearly sensitive to
the qubit frequency detuning, resulting in a gate error that is quadratically sensitive
to the qubit frequency detuning.
At a one-percent
qubit frequency detuning ($\delta f_{q} / f_{q} = 1\%$)
the analytic gate achieves a gate error $\sim 4.5 \cdot 10^{-5}$,
which is sufficient for quantum error correction.
Although the analytic $Z/2$ gate performs well, it
only works at the gate time $t_{Z/2}$. The ability to perform $Z$
rotations in arbitrary times is critical
for multi-qubit experiments, where the qubits operate at different
frequencies $f_{q, i} \neq f_{q, j}$.
Each numerical method can find solutions at
all gate times above $t_{Z/2}$, see Figure \ref{fig:staticb}.
These numerical methods offer an effective scheme for synchronizing
multi-qubit experiments.

The sampling and unscented sampling methods combine idling periods
with time-anti-symmetric ramps.
The gate error at $\delta f_{q} / f_{q} = 1\%$ for the sampling method
does not improve substantially over the
range of gate times. The unscented sampling method
achieves linear decreases in its gate error at
$\delta f_{q} / f_{q} = 1\%$ with longer gate times
until half the Larmor period $1 / 2 f_{q}$ after which it achieves a consistent
gate error $\sim 3.5 \cdot 10^{-5}$.

The derivative methods converge on qualitatively similar solutions that
use fast triangle pulses at the boundaries and balance time
on either side of the flux-frustration point symmetrically at low amplitudes.
Both methods achieve a super-linear scaling in their gate error as
a function of their gate time.
We believe the 1\textsuperscript{st}-order method outperforms the 2\textsuperscript{nd}-order
method due to the low contribution of second-order
terms to the gate error in this deviation regime, see Appendix \ref{appendix:derivative}.
The gate error for the 1\textsuperscript{st}-order
derivative method approaches zero at the Larmor period $1 / f_{q}$, see Figure \ref{fig:staticc}.
This result mimics the
ability of composite pulses to mitigate parameter deviation errors to arbitrary
order with sufficiently many pulses \cite{merrill2014progress}.
It is difficult to choose an appropriate composite pulse
for the problem studied here due to our Hamiltonian and experimental constraints.
We propose comparisons between composite pulses and numerical techniques
for future work.

%% F3
\begin{figure*}[ht]
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f3a.png}
    \caption{}
    \label{fig:stochastica}
  \end{subfigure}\hspace{0.05\textwidth}
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f3b.png}
        \caption{}
    \label{fig:stochasticb}
  \end{subfigure}
  \caption{
    (a) $X/2$ gates robust to flux amplitude offsets constructed with the analytic,
    sampling, unscented sampling, and the 1\textsuperscript{st}-
    and 2\textsuperscript{nd}-order derivative methods. The gates shown
    are the solutions at the analytic gate time.
    (b) Simulation of stochastic 1/$f$ flux noise for
    successive gate applications. The cumulative
    gate error is computed after each gate application.
  }
  \label{fig:stochastic}
\end{figure*}

### 2020-11-30 paper/6stochastic.tex
updated figures, lots of proofreading

\section{Robustness to Time-Dependent Parameter Deviations \label{sec:stochastic}}
An additional source of experimental error arises from time-dependent
Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
superconducting circuit elements, magnetic flux noise is a significant
source of coherent errors. Magnetic flux noise
modifies the fluxonium Hamiltonian \eqref{eq:hamiltonian}
by $a(t) \gets a(t) + \delta a(t)$.
It is well studied that the spectral density of $\delta a(t)$ follows a
1/$f$ distribution for a range of devices, consisting primarily of low frequency
noise \todo{1/$f$ citations}. Analytic methods to combat flux noise
take advantage of the low frequency characteristic and
treat the noise as quasi-static, performing generalizations of the spin-echo technique
to compensate for erroneous drift. This is the strategy employed by the analytic gate
considered here.

We task the analytic gate and those produced by
the robust control methods of the previous section
to realize an $X/2$ gate subject to 1/$f$ flux noise.
We compute gate errors for each method
by evolving the optimized trajectory under the Hamiltonian in \eqref{eq:hamiltonain}
with $a(t) \gets a(t) + \delta a(t)$.
The flux noise $\delta a(t)$ is generated by
filtering white noise sampled from a standard normal distribution with a finite
impulse response filter \cite{saspweb2011}.
It is then scaled by the 
flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
\delta a (t) \sim \sigma_{a} = 2.5 \cdot 10^{-5} \textrm{GHz}$.
The unscented sampling method is modified so that its samples
are subject to 1/$f$ flux noise by carrying the state of a finite impulse response filter
in the augmented state vector \eqref{eq:astatecontrols}.
In principle the sampling method could be modified
similarly but we choose to sample statically at $\sigma_{a}$ for comparision.
The derivative methods require no algorithmic modification from the static case,
but the TDSE is now differentiated with respect to $a(t)$ instead of $f_{q}$ as
in \eqref{d1dyn}.

We simulate successive applications of the gate constructed by each method
and compute the cumulative gate error
after each application, see Figure \ref{fig:stochastic}. Both the analytic
and numerical methods achieve single gate errors
sufficient for quantum error correction.
Despite converging on qualitatively different solutions, the
numerical methods perform similarly in the concatenated
gate application comparision. They achieve a two
order of magnitude cumulative gate error reduction over the analytic method after $200$
gate applications $\sim 11 \mu\textrm{s}$.
1/$f$ flux noise is a signficiant source of coherent errors in NISQ applications and
these numerical techniques offer effective avenues to mitigate them.

>>>>

\section{Robustness to Time-Dependent Parameter Deviations \label{sec:stochastic}}
An additional source of experimental error arises from time-dependent
Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
superconducting circuit elements, magnetic flux noise is a significant
source of coherent errors. Magnetic flux noise
modifies the fluxonium Hamiltonian \eqref{eq:hamiltonian}
by $a(t) \gets a(t) + \delta a(t)$.
The spectral density of $\delta a(t)$ follows a
1/$f$ distribution for a range of devices, consisting primarily of low frequency
noise \cite{bialczak20071f, koch2007model,
  yoshihara2006decoherence, yoshihara2010correlated}.
Analytic methods to combat flux noise
take advantage of the low-frequency characteristic and
treat the noise as quasi-static, performing generalizations of the spin-echo technique
to compensate for erroneous drift. The analytic gate considered here
follows this strategy.

We task the analytic gate and those produced by
the robust control methods of the previous section
to realize an $X/2$ gate subject to 1/$f$ flux noise.
We compute gate errors for each method
by evolving system under the Hamiltonian in \eqref{eq:hamiltonian}
where the optimized flux amplitude is modified $a(t) \gets a(t) + \delta a(t)$.
The flux noise $\delta a(t)$ is generated by
filtering white noise sampled from a standard normal distribution with a finite
impulse response filter \cite{saspweb2011}.
It is then scaled by the 
flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
\delta a (t) \sim \sigma_{a} = 2.5 \cdot 10^{-5} \textrm{GHz}$.
The unscented sampling method is modified so that its samples
are subject to 1/$f$ flux noise by carrying the state of a finite impulse response filter
in the augmented state vector \eqref{eq:astatecontrols}.
In principle, we could modify the sampling method
similarly but we choose to sample statically at $\sigma_{a}$ for comparison.
The derivative methods require no algorithmic modification from the static case,
but the TDSE is now differentiated with respect to $a(t)$ instead of $f_{q}$ as
in \eqref{eq:d1dyn}.

We simulate successive applications of the gate constructed by each method
and compute the cumulative gate error
after each application, see Figure \ref{fig:stochastic}. Both the analytic
and numerical methods achieve single gate errors
sufficient for quantum error correction.
Despite converging on qualitatively different solutions, the
numerical methods perform similarly in the concatenated
gate application comparison. They achieve a two
order of magnitude cumulative gate error reduction over the analytic method after $200$
gate applications $\sim 11 \mu\textrm{s}$.
1/$f$ flux noise is a significiant source of coherent errors in NISQ applications and
these numerical techniques offer effective avenues to mitigate them.
